---
title: "p8105_hw3_rc3520"
author: "Ruyi Chen"
date: "2023-10-10"
output: github_document
---

```{r setup, include=FALSE}
library (tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets)
```

## R Markdown
#Problem 1

##load the data from p8105.datasets
```{r}
data("instacart")
instacart_df = 
  instacart |> 
  as_tibble()
```

## Look at data summary
```{r eval = FALSE}
str(instacart_df)
skimr::skim(instacart_df)
```

## Data Description:
The dataset 'instacart' includes `r nrow(instacart_df)` observations and `r ncol(instacart_df)` variables. Every row includes information representing a single product from an order on instacart. Identifiers for order include variables describing date, time and number of days since prior order; identifiers for product indlude variables describing product name, such as Bulgarian Yogurt and Cucumber Kirby, department, such as dairy eggs and produce, aisle, such as yogurt and fresh vegetables and add-to-cart order.In total, there are `r instacart_df |> select(product_id) |> distinct() |> count()` products found in `r instacart_df |> select(user_id, order_id) |> distinct() |> count()` orders from `r instacart_df |> select(user_id) |> distinct() |> count()` distinct users.

## Answer the following questions
### 1. There are 134 aisles; most items are ordered from fresh vegetables and fresh fruits.
```{r}
instacart_df |>
  count(aisle) |>
  arrange(desc(n))
```


### 2. Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r}
instacart_df |> 
  count(aisle) |> 
  filter(n > 10000) |> 
  mutate(aisle = fct_reorder(aisle, n)) |> 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

### 3. Our next table shows the three most popular items in aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits`, and includes the number of times each item is ordered in your table.
```{r}
instacart_df |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle) |> 
  count(product_name) |> 
  mutate(rank = min_rank(desc(n))) |> 
  filter(rank < 4) |> 
  arrange(desc(n)) |>
  knitr::kable()
```

Finally is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. This table has been formatted in an untidy manner for human readers. Pink Lady Apples are generally purchased slightly earlier in the day than Coffee Ice Cream, with the exception of day 5.
```{r}
instacart_df |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarize(mean_hr = mean(order_hour_of_day)) |>
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hr) |>
  knitr::kable(digits=2)
  
```



# Problem 2
##load the data from p8105.datasets
```{r}
data("brfss_smart2010")

brfss_smart2010 = 
  brfss_smart2010 |> 
  as_tibble()
brfss_smart2010

```

## Data Cleaning
```{r}
brfss =
  brfss_smart2010 |> 
  janitor::clean_names() |>
  filter(topic %in% c("Overall Health")) |>
  mutate(response = factor(response, levels = c("Excellent","Very good","Good", "Fair", "Poor")))
view(brfss)
```
## Answer Questions
### 1.In 2002, CT, FL, MA, NC, NJ, and PA were observed at 7 or more locations.
```{r}
brfss |>
  filter(year %in% c(2002))|>
  group_by(locationabbr)|>
  summarise(count_locations = n_distinct(locationdesc)) |>
  filter(count_locations >= 7)
```
### 2. In 2010, CA, CO, FL, MA, MD, NC, NE, NJ, NY, OH, PA, SC, TX, and WA were observed at 7 or more locations.
```{r}
brfss |>
  filter(year %in% c(2010))|>
  group_by(locationabbr)|>
  summarise(count_locations = n_distinct(locationdesc)) |>
  filter(count_locations >= 7)
```
### 3. Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).
```{r}
# Construct a dataset
brfss_excellent =
  brfss |>
  filter(response %in% c("Excellent"))|>
  group_by (locationabbr, year)|>
  summarise( ave_data_value = mean(data_value, na.rm = TRUE))
```

```{r}
#Make a plot
ggp_brfss_excellent =
  brfss_excellent |>
  ggplot(aes(x = year, y = ave_data_value, group = locationabbr, color = locationabbr)) + 
  geom_point() + 
  geom_line() +
  labs(title = "Data value average over time within each state",
       x = "Year", y = "Average Crude Score", color = "State") +
  theme_minimal() +
  theme(legend.position = "bottom")
ggp_brfss_excellent
```

### 4. Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.
```{r}
ggp_brfss_response = 
  brfss |>
  filter (year %in% c(2006,2010), locationabbr %in% c("NY")) |>
  group_by (year) |>
  ggplot(aes(x= response, y= data_value, fill = response))+
  geom_boxplot() + 
  facet_grid(. ~ year)+
  labs(title = "Distribution of data_value for response in NY for 2006 and 2010",
       x = "Response", y = "Crude Score") +
  theme_minimal() 
ggp_brfss_response
```

